Discussion

### 2021-07-04 text/discussion.Rmd
Update text for cluster 7 and 8 in discussion

Our results illustrate how unique combinations of mutations in pan-cancer driver
genes with tissue-specific pathway disruptions result in common categories when
viewed at the level of the pathway knowledge-base. Top cancer driver genes
(_e.g. PIK3CA_ and _TP53_) are found in most of the clusters, in spite of the
fact that they contribute to many cluster-specific pathways. This can only be
explained by less commonly mutated genes complementing the unique combination of driver
genes in each tumor, and we speculate that many of these less commonly mutated genes are
sensitive to increased mutation frequency in different tissues. For example, the most 
frequently mutated genes (_TP53_, _PIK3CA_ and _MUC16_) in cluster 2 were also frequently 
mutated in several other clusters. Yet, pathway enrichment in cluster 2 was defined 
by less commonly mutated genes in the proteasomal degradation pathway (_e.g. PSMA3_, 
_PSMA1_ or _PSME4_). The ubiquitin proteasome system maintains cellular protein 
homeostasis by regulating protein turnover. Pathway mutations found in cluster 2 
(_e.g._ “B-catenin degradation”, “cellular response to hypoxia”, “regulation of cell cycle” 
and _etc_) covered several aspects of the cancer hallmarks described by Hanahan 
and Weinberg. These hallmarks included “sustaining proliferative signaling”, “evading 
tumor suppressors” and “resisting cell death” [@Bhattacharjee2014]. Cluster 
3 was defined by less commonly mutated genes related to ribosomal subunit. Pathway 
enrichment for cluster 3 included dysregulation in “rRNA processing in the nucleus 
and cytosol” and consequently, pathways involved in eukaryotic translation and 
nonsense mediated decay. Mutations in the ribosomal subunit have been associated 
with selective translation of mRNA to upregulate proliferation [@kampen2019]. 
Pathway enrichment in cluster 3 consist of hallmarks relating to “sustaining proliferative 
signaling”, “evading tumor suppressors”, “resisting cell death”, “deregulating 
cellular energetics”, and “genome instability and mutations”. Cluster 4 was defined 
by less commonly mutated genes related to nucleosome structure that resulted in 
pathway enrichment in DNA methylation and DNA double-strand break repair pathway. 
Cluster 4 was also enriched for “telomere maintenance”, a mechanism contributing to 
cancer progression by counteracting telomere shortening. Mutations in the 
nucleosome as found in cluster 4 have been proposed to hit all attributes of hallmarks of cancer. Cluster 
7 and 8 shared multiple enrichment in signaling pathways related to the hallmarks: 
“sustaining proliferative signaling”, “evading tumor suppressors”, “resisting cell 
death”, and “inducing angiogenesis”. These clusters were strongly defined by 
mutations in _PI3K_ or _RAS_, respectively. Taken together, our clusters highlight 
the similarities and differences at a gene and pathway level between and within 
different cancer types – uncovering the molecular mechanism in cancer. In the future,
we hope to explore the defining features of the clusters by connecting the pathway 
mutations to cancer hallmarks.

>>>>

Our results illustrate how unique combinations of mutations in pan-cancer driver
genes with tissue-specific pathway disruptions result in common categories when
viewed at the level of the pathway knowledge-base. Top cancer driver genes
(_e.g. PIK3CA_ and _TP53_) are found in most of the clusters, in spite of the
fact that they contribute to many cluster-specific pathways. This can only be
explained by less commonly mutated genes complementing the unique combination of driver
genes in each tumor, and we speculate that many of these less commonly mutated genes are
sensitive to increased mutation frequency in different tissues. For example, the most 
frequently mutated genes (_TP53_, _PIK3CA_ and _MUC16_) in cluster 2 were also frequently 
mutated in several other clusters. Yet, pathway enrichment in cluster 2 was defined 
by less commonly mutated genes in the proteasomal degradation pathway (_e.g. PSMA3_, 
_PSMA1_ or _PSME4_). The ubiquitin proteasome system maintains cellular protein 
homeostasis by regulating protein turnover. Pathway mutations found in cluster 2 
(_e.g._ “B-catenin degradation”, “cellular response to hypoxia”, “regulation of cell cycle” 
and _etc_) covered several aspects of the cancer hallmarks described by Hanahan 
and Weinberg. These hallmarks included “sustaining proliferative signaling”, “evading 
tumor suppressors” and “resisting cell death” [@Bhattacharjee2014]. Cluster 
3 was defined by less commonly mutated genes related to ribosomal subunit. Pathway 
enrichment for cluster 3 included dysregulation in “rRNA processing in the nucleus 
and cytosol” and consequently, pathways involved in eukaryotic translation and 
nonsense mediated decay. Mutations in the ribosomal subunit have been associated 
with selective translation of mRNA to upregulate proliferation [@kampen2019]. 
Pathway enrichment in cluster 3 consist of hallmarks relating to “sustaining proliferative 
signaling”, “evading tumor suppressors”, “resisting cell death”, “deregulating 
cellular energetics”, and “genome instability and mutations”. Cluster 4 was defined 
by less commonly mutated genes related to nucleosome structure that resulted in 
pathway enrichment in DNA methylation and DNA double-strand break repair pathway. 
Cluster 4 was also enriched for “telomere maintenance”, a mechanism contributing to 
cancer progression by counteracting telomere shortening. Mutations in the 
nucleosome as found in cluster 4 have been proposed to hit all attributes of hallmarks of cancer. Cluster 
7 and 8 shared multiple enrichment in signaling pathways related to the hallmarks, 
showing that some clusters share common driver mechanisms: 
“sustaining proliferative signaling”, “evading tumor suppressors”, “resisting cell 
death”, and “inducing angiogenesis”. These clusters were strongly defined by 
mutations in _PI3K_ or _RAS_, respectively, suggesting a shared etiology in the EGFR 
pathway for which these genes represent alternative intracellular signaling mechanisms.Taken together, our clusters highlight 
the similarities and differences at a gene and pathway level between and within 
different cancer types – uncovering the molecular mechanism in cancer. In the future,
we hope to explore the defining features of the clusters by connecting the pathway 
mutations to cancer hallmarks.
